The future age structure of scientific disciplines

### 2025-04-12 age_structure_forecasts.qmd
Started on quarto doc

```{r setup, include=FALSE}
library(ggplot2)
library(patchwork)
library(vital)
library(fable)
library(dplyr)
library(tidyr)
library(stringr)
library(forcats)
library(knitr)
library(purrr)
library(kableExtra)
library(gganimate)
library(ggokabeito)

@@

```{r}
#| label: fig-labour-force
#| fig-cap: Labour force participation rates by one-year age groups for 2016 and 2021, along with the average of these two years, which is used as a proxy for 2006 and 2011.

census_science <- census |>
  filter(category == "Natural and Physical Sciences")

cols <- c("#FF6F00", "#009E73", "#009DFF", "#FF00BF", "#000000")
names(cols) <- c(unique(census_science$Year), "Average")

census_science <- census_science |>
  mutate(
    age_group = fct_relevel(age_group, "100+", after = Inf),
    Year = factor(Year, levels = c(2006, 2011, 2016, 2021))
  )

census_participation <- census_science |>
  mutate(
    Year = case_when(
      Year == 2011 ~ "Average",
      TRUE ~ as.character(Year)
    ),
    age_group = fct_relevel(age_group, "100+", after = Inf),
    Year = factor(Year, levels = c(2006, "Average", 2016, 2021))
  )

ggplot(census_participation |> filter(Year %in% c("Average", 2016, 2021)), 
       aes(x = as.numeric(gsub("\\+", "", age_group)), y = participation, 
           shape = as.factor(Year), color = as.factor(Year))) +
  geom_point(size = 1.5) +
  labs(
    title = "Labour Force Participation",
    subtitle = "Natural and Physical Sciences",
    x = "One-year age groups",
    y = "Participation rate",
    shape = "Census Year",
    color = "Census Year"
  ) +
  scale_x_continuous(breaks = seq(20, 100, by = 10), labels = c(seq(20, 90, by = 10), "100+")) +
  scale_shape_manual(values = c(1, 2, 0)) +
  scale_color_manual(values = cols) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

@@

```{r}
#| label: fig-census
#| fig-cap: Distribution of scientists active in the labour market by age across Census years (2006, 2011, 2016, 2021).
years <- seq(2006, 2021, by = 5)
cols <- c("#FF6F00", "#009E73", "#009DFF", "#FF00BF")
names(cols) <- years
natural_and_physical_sciences |>
  as_tibble() |>
  filter(Year %in% years) |>
  mutate(Year = factor(Year, levels = rev(years))) |>
  ggplot() +
  aes(x = Age, y = Working, color = Year, group = Year) +
  geom_line() +
  labs(y = "Number of active scientists") +
  scale_color_manual(values = cols, name = "Census Year") +
  labs(
    title = "Working Population",
    subtitle = "Natural and Physical Sciences"
  ) +
  scale_x_continuous(breaks = seq(20, 100, by = 10)) +
  theme_classic()  +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

```{r}
#| label: fig-cohort-interpolation
#| fig-cap: Interpolated number of scientists in Australia, 2006–2021, using linear cohort interpolation.

natural_and_physical_sciences  |>
  autoplot(Working) +
  labs(y = "Number of active scientists") +
  labs(
    title = "Interpolated Working Population",
    subtitle = "Natural and Physical Sciences\n2006 - 2021"
  ) +
  scale_x_continuous(breaks = seq(20, 100, by = 10)) +
  theme_classic()  +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

@@

```{r}
#| label: fig-beforedisaggregation
#| fig-cap: Retirement intentions for key industries before smoothed disaggregation.

industry_retirement <- bind_rows(
  retirement_data |> mutate(industry = "Professional, scientific and technical services"),
  industry_retirement  
)

ggplot(industry_retirement, aes(x = age_group, y = pc)) +
  geom_col(alpha=0.6) +
  facet_wrap(~industry, ncol=2) +
  labs(
    title = "Retirement Intentions Across Key Industries",
    subtitle = "2022–23 Financial Year",
    x = "Age group",
    y = "Percentage of retirement intentions"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
```

```{r}
#| label: fig-afterdisaggregation
#| fig-cap: Retirement intentions across key industries after smoothed disaggregation.

all_retirement |> 
  ggplot(aes(x = Age, y = pc, group = industry)) +
  geom_line() +
  facet_wrap(~industry, ncol=2) +
  labs(
    y = "Percentage of retirement intentions",
    title = "Retirement Intentions Across Key Industries",
    subtitle = "2022–23 Financial Year"
  ) +
  scale_x_continuous(breaks = seq(40, 100, by = 10)) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

>>>>

```{r setup, include=FALSE}
#| eval: true
library(ggplot2)
library(patchwork)
library(knitr)
library(kableExtra)

@@

```{r}
#| label: fig-labour-force
#| fig-cap: Labour force participation rates by one-year age groups for 2016 and 2021, along with the average of these two years, which is used as a proxy for 2006 and 2011.
tar_read(fig1)
```

@@

```{r}
#| label: fig-census
#| fig-cap: Distribution of scientists active in the labour market by age across  Census years (2006, 2011, 2016, 2021).
tar_read(fig2)
```

```{r}
#| label: fig-cohort-interpolation
#| fig-cap: Interpolated number of scientists in Australia, 2006–2021, using linear cohort interpolation.
tar_read(fig3)
```

@@

```{r}
#| label: fig-beforedisaggregation
#| fig-cap: Retirement intentions for key industries before smoothed disaggregation.
tar_read(fig4)
```

```{r}
#| label: fig-afterdisaggregation
#| fig-cap: Retirement intentions across key industries after smoothed disaggregation.
tar_read(fig5)
```
